\chapter{Experiment 2 - In-Network Malware Dataset Transformation}
\label{chap:four}

Here, our results are divided into two sections. 
The first considers the efficacy of the malware classifier and where certain improvements could be made.
The second section considers how the internal and external representation of our datasets influenced the ability of the neural network to learn.

\sect{Malware Classifier} \label{malware classifier}
As shown in Table~\ref{Tab:test}, the Random forest classifier outperforms all of the neural network and support vector machine architectures.
It is also noteworthy that the Fourier-transformed data fed to the convolutional Neural Network appears to misclassify an extremely high percentage of samples, and by flipping the labels on the test samples, we actually achieve 72.6\% accuracy, making it our second most accurate neural network classifier.

 
\begin{table}[ht]
\caption{Android App Data Test Results}
\centering
\label{Tab:test}	
\begin{tabular}{lll}
Data and Architecture Combination & Test Accuracy & mean step time (ms) \\
Raw, Fully-Connected NN           & 62.60\%         & 13                  \\
Raw, Convolutional NN             & 72.89\%         & 54                  \\
Fourier, Fully-Connected NN       & 58.88\%         & 12                  \\
Fourier, Convolutional NN         & 70.81\%         & 56                  \\
Wavelet, Fully-Connected NN       & 61.95\%         & 12                  \\
Wavelet, Convolutional NN         & 70.77\%         & 52                  \\
Raw, Fourier NN                   & 66.59\%         & 553                 \\
Raw, Wavelet NN                   & 73.33\%         & 623                \\
Raw, Random Forest                & 79.80\%         & N/A                 \\
Raw, Support Vector Classifier    & 55.28\%         & N/A                
\end{tabular}
\end{table}
 
\sect{Data Representation and Neural Networks} \label{data representation}
We explored 3 different data representations: Raw, Fourier-transformed, and Wavelet-transformed.
The summary statistics of this data are displayed in Table~\ref{Tab:summary}.
The results achieved by the classifiers as displayed in Table~\ref{Tab:test} suggest the following:
\begin{enumerate}
\item Both the Fully-Connected and Convolutional neural networks have difficulty learning from the Fourier transformed data
\item The wavelet data and the raw data result in comparable performance across both fully-connected and convolutional neural networks.
\end{enumerate}

The inability to learn from the Fourier-transformed data may be due to the variance of that dataset being four orders of magnitude larger than the raw data.
Given the tremendous variance even within a single sample type, the error which gets backpropogated through the networks is almost certainly going to be far too noisy to train the weights such that they can appropriately predict future samples. 
